### 2016-12-06 poster/poster.tex
added poster file and some graphics

\documentclass[paperwidth=118cm,paperheight=84cm,portrait,margin=8em,fontscale=0.3]{baposter}
%\usepackage{acl-hlt2011}
%84 cm wide x 118 cm high
\usepackage{times}
\usepackage{svg}
\usepackage{color}
\usepackage{multirow}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{setspace}

%%30in X 40in
%available sizes: (24 X 38?) (36X48)
%\usepackage{array}
\usepackage{relsize}		% For \smaller
\usepackage{url}			% For \url
\usepackage{epstopdf}	% Included EPS files automatically converted to PDF to include with pdflatex

\newcommand{\compresslist}{
	\setlength{\itemsep}{1pt}
	\setlength{\parskip}{0pt}
	\setlength{\parsep}{0pt}
}
\newenvironment{itquote}
{\begin{quote}\begin{itshape}\begin{small}	\begin{center}\setlength{\parskip}{0.0em}\setlength{\parsep}{0.0em}}
{\end{center}\end{small}\end{itshape}\end{quote}}

\graphicspath{{./}}	% Root directory of the pictures 
\tracingstats=2			% Enabled LaTeX logging with conditionals
%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{uwPurple}{HTML}{4b2e83}
%\definecolor{uwPurple}{HTML}{3b3e72}
\definecolor{uwGold}{RGB}{232,211,162}
\definecolor{uwMetallicGold}{RGB}{145,123,76}

\begin{document}
\typeout{Poster rendering started}

%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\background{
}

\begin{poster}{
	%grid=true,
	columns=3,
	colspacing=2em,
	headerheight = 0.17\textheight ,
	%eyecatcher=false, 
	borderColor=uwMetallicGold,
	headerColorOne=uwPurple,
	headerColorTwo=uwPurple,
	headerFontColor=white,
	% Only simple background color used, no shading, so boxColorTwo isn't necessary
	boxColorOne=white,
	boxColorTwo=uwPurple,
	headerborder=closed,
	headershape=rounded,
	headerfont=\Large\sf\bf,
	textborder=none,
	bgColorOne=uwGold,
	bgColorTwo=uwPurple,
	background=plain,
	headerborder=open,
	boxshade=plain
}
{
%Eye Catcher, empty if option eyecatcher=no
{
	\begin{tabular}{l}
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
			\includegraphics[width=2.5in]{graphics/UWlogo.pdf} \\
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
	\end{tabular}
	%\includegraphics[scale=1.7]{graphics/pnt5in.png}
}
}
{ % GUIDE: GUI for Document Exploration
\textbf{$\spadesuit$ SPADE -- SPan Annotated Document Exploration $\spadesuit$}
}
{
\vspace*{0.8em}
\textbf{Lucy Lin, George Mulcaire \& Maarten Sap}\\

%\vspace*{0.5em}
CSE 512 Data Visualization Final Project\\
Computer Science and Engineering, University of Washington

 \vspace*{6pt}
}
{
	\begin{tabular}{r}
			\includegraphics[width=1.80in]{graphics/CSElogo.png}\includegraphics[scale=.7]{graphics/pnt5in.png}
	\end{tabular}
}

%For poster contributors, the format is as follows.  Posterboards are 3 feet (90cm) high, 4 feet (120cm) wide self-standing boards that will be placed on top of conference tables. Materials for attaching posters to the boards (double-sided tape, pushpins, clips) will be provided.  Poster authors should finish setting up their posters by the end of the morning break, at the latest.  Also, please note the short "Poster Teasers" in the schedule before the poster session, where you will have 60 seconds (strictly timed!) to give a quick overview that will get people excited about coming to see your poster.
%
%For everyone, please bear in mind that a central goal of the workshop is to create discussion between computational linguists and clinical psychologists.  Therefore we would ask that you please formulate your presentations or posters for a diverse audience, rather than the more typical, purely technical ACL crowd.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Column 0

%% Box 0
\headerbox{Deep Learning and NLP}{name=intro1,column=0,row=0.00, span=1}{
\color{black}
\vspace*{5pt}
\textbf{NLP for NLP}
\begin{itemize}\addtolength{\itemsep}{-0.25\baselineskip}
\item nlp is natural
\item the natural is about the processing, not the language
\end{itemize}\vspace*{5pt}

\begin{tabular}{cc}
\includegraphics[width=.24\columnwidth]{graphics/python.png}
\hspace*{20pt} & \hspace*{-15pt}\parbox{0.62\textwidth}{
\vspace*{-40pt}\textbf{Deep learning}
\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
\item wow
\item neurons
\item just a function
\item such deep
\vspace*{-5pt}
\end{itemize}}
\end{tabular}
}

%% Box 2
\headerbox{Dataset}{name=intro1,column=0,row=0.40, span=1}{
\color{black}
English GigaWord corpus (LDC), LA Times from 2009 ? \vspace*{10pt}
\\
\begin{tabular}{cc}
\parbox{0.45\textwidth}{
\vspace*{-0pt}
\includegraphics[width=.24\columnwidth]{graphics/newspaper.pdf}
\includegraphics[width=.24\columnwidth]{graphics/LAtimes.jpeg}

}
&
\hspace*{-10pt}
\parbox{0.50\textwidth}{
\textbf{Data}
\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
\item Split into sentences then words using NLTK
\item $1,191,848$ sentences
\item $27,269,856$ words (``tokens'') total
\item $35,642$ distinct words in the end
%\vspace*{-5pt}
\end{itemize}}
\end{tabular}
}

%% Box 2
\headerbox{Hyper params?}{name=intro1,column=0,row=0.75, span=1}{
\color{black}
??
\vspace{.4pt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLUMN 1

%% BOX 0
\headerbox{Logos \& Graphs}{name=results,column=1,row=0.00, span=2}{
\includegraphics[width=.24\columnwidth]{graphics/TensorFlow.png}
\includegraphics[width=.24\columnwidth]{graphics/Theano.jpeg}
\includegraphics[width=.24\columnwidth]{graphics/DyNet.pdf}
}

%% BOX 1
%\headerbox{Differential Word Clouds}{name=results,column=1,row=0.45, span=1}{
%\begin{center}
%\includegraphics[scale=.5]{graphics/wordcloud1.pdf}
%\includegraphics[scale=.5]{graphics/wordcloud2.pdf}
%\end{center}
%
%Something about the method for creating them ?
%}

%% Box 3
\headerbox{What else?}{name=methods,column=1,row=0.69, span=1}{
\color{black}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLUMN 2

%% BOX 0
%\headerbox{Search Results}{name=results,column=2,row=0, span=1}{
%\begin{center}
%\includegraphics[scale=.5]{graphics/searchResults.JPG}
%
%\end{center}
%\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}%\addtolength{\itemindent}{-5pt}
%\item Search for words or phrases in the document text
%\item Filter on frames to examine relationships between the language used and the frame evoked
%\item Search via the wordcloud, the search bar, or by clicking on frames in search results
%\end{itemize}
%}

%% BOX 1
%\headerbox{Document Viewer}{name=results,column=2,row=0.47, span=1}{
%\begin{center}
%\includegraphics[scale=.4]{graphics/documentView.png}
%\end{center}
%
%The document viewer enables a user to examine a specific text with
%clearly marked frame annotations.
%}

%% BOX 2
\headerbox{How many GPUs should I buy?}{name=results,column=2,row=0.69, span=1}{
}

\end{poster}

\end{document}