### 2016-12-14 paper/main.tex
paper proofread

\documentclass{article}
\usepackage[utf8]{inputenc}

% align enviroment with reduced spacing around it
\usepackage{amsmath}
\usepackage{etoolbox}
\newcommand{\zerodisplayskips}{%
  \setlength{\abovedisplayskip}{0.25em}
  \setlength{\belowdisplayskip}{0.25em}
  \setlength{\abovedisplayshortskip}{0.25em}
  \setlength{\belowdisplayshortskip}{0.25em}}
\appto{\normalsize}{\zerodisplayskips}

\usepackage{emnlp16}
\usepackage{graphicx,hyperref}
\usepackage{tabularx}
\usepackage[table,xcdraw]{xcolor}

\newcommand{\ms}[1]{{\color{cyan}\{\textit{#1}\}$_{ms}$}}
\newcommand{\lhl}[1]{{\color{magenta}\{\textit{#1}\}$_{lhl}$}}

\title{Benchmarking Python Deep Learning Frameworks\\ for Language Modeling on GPUs}
% feel free to make this title much much better!
\emnlpfinalcopy

\author{Lucy Lin, George Mulcaire \& Maarten Sap
\\University of Washington}
\date{December 2016}

\begin{document}

\maketitle

\begin{abstract}
Neural networks are omnipresent in natural language processing (NLP).
We benchmark three popular Python frameworks (DyNet, TensorFlow, and Theano) on the standard NLP task of language modeling, and find that DyNet is significantly faster on this task. We also discuss other bottlenecks beyond performance, such as ease of use, that may impact the selection of a neural network framework.
\end{abstract}

\section{Introduction}
In recent years, deep learning approaches have exploded in popularity due to their performance gains over traditional systems on many tasks. Several deep learning frameworks have been released in response to this, each claiming niche improvements over others, but it is not necessarily clear which frameworks are better suited for which types of applications. Previous comparisons betewen frameworks may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).

In this project, we benchmark deep learning frameworks on a complex natural language processing (NLP) task and rank them based on different performance metrics. Specifically, we look at the following frameworks:
\begin{itemize}
	\item TensorFlow \cite{tensorflow}
	\item Theano \cite{theano}, using the Keras wrapper \cite{keras}
	\item DyNet \cite{dynet}
\end{itemize}
In NLP, recurrent neural networks (RNNs) are of particular interest because they make use of the sequential nature of language. Benchmarks on RNNs have been run before, but most tasks were benchmarked on toy tasks which do not necessarily reflect performance on real data. For instance, GitHub user \verb!glample! generated random data to run through their RNN \cite{glample}, which is not realistic. A standard approach in NLP is to transform words into vocabulary-sized one-hot vectors and then embed them as hidden-sized vectors (i.e., dimensionality reduction). Sparse learning of the embedding matrix, which is usually in the order of $[20,000 \times 200]$, is likely to severely impact performance.

\section{Background}
\subsection{Language Modeling}
The task of language modeling is about estimating probabilities of particular sequences of words. It is used as an invaluable component in various real-world applications such as speech recognition and statistical machine translation. More formally, if $s=\{START,x_1,x_2,...,x_n,STOP\}$ is an n-word sentence, language modelling will try to estimate
the probability of each word $x_i$ given its history: $p(x_i|START,x_1,x_2,...,x_{i-1})$.

Historically, this task was done by simply using an l-th order Markov model:
\begin{align*}
    p(x_i|START,x_1,x_2,...,x_{i-1}) = \\
        p(x_i|x_{i-l},x_{i-l+1},...,x_{i-1})
\end{align*}
However, the Markov assumption does not need to be made with RNNs, as they are capable of encoding longer histories. With RNNs, we normalize (using softmax) the output of a nonlinear function $f$:
\begin{align*}
	f\left(p(x_i|START,x_1,x_2,...,x_{i-1})\right) = \\
		softmax\left(f(x_i,START,x_1,x_2,...,x_{i-1})\right)
\end{align*}

\subsection{Frameworks}
All three frameworks use a symbolic computation graph, which maps operations and variables to nodes in a graph. To perform a computation, one has to ``query'' the graph by providing values for variables needed to compute the desired output.
Figure~\ref{fig:compGraph} shows an example of a simple computation graph.

\begin{figure}\begin{center}
\includegraphics[scale=.4]{graphics/tensorflowGraph.png}
\caption{\label{fig:compGraph}Simple neural computation graph example. (\texttt{ReLU} is a rectified linear unit.) Figure is taken from \protect\cite{tensorflow}.} 
\end{center}\end{figure}

The frameworks handle computation graphs differently. Theano and Tensorflow perform ``static'' declaration of computation graphs: first, the graph architecture is specified, and then data is run through the compiled graph to train or make predictions. In contrast, DyNet does ``dynamic'' declaration of computation graphs; that is, the computation graph is defined on-the-fly as operations are executed.

All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper seems to most reliably support Python 2. TensorFlow and Theano are primarily Python frameworks that support both Python 2 and 3.} that wrap C++ or optimized Python code (e.g., numpy).

For speed, the frameworks all use fast low-level linear algebra libraries for matrix and tensor operations. Both TensorFlow and DyNet make extensive use of the Eigen library for C++, while Theano uses BLAS instead.

%TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.

\section{Experiments}
\subsection{Data \& Preprocessing}
We chose to focus on the Los Angeles Times subset from 2009 of the GigaWord corpus \cite{gigaword}. Each of these documents is a news article, so we used NLTK’s \cite{nltk} sentence splitter on the articles. Every sentence is then tokenized using NLTK’s \verb!casual_tokenize! function, limiting ourselves to sentences that are $<100$ tokens long. In order to limit overfitting, we replace all words that occurred less than $150$ times by a special \verb!OOV! symbol (this allows the model to be more robust when encountering previously unseen words). We ended up with $1,191,848$ sentences, $27,269,856$ total word tokens and a vocabulary size of $35,642$.

\subsection{Implementation \& Hyperparameters}
We implemented the same RNN language model in each framework. The model consists of a matrix of input word embeddings, which are passed through a RNN to get predicted word embeddings and then compared (with a loss function) to a corresponding set of output embeddings. All these elements are available in all frameworks (in Theano's case, the RNN is provided by the Keras wrapper). Our implementations are made available on GitHub.\footnote{\url{https://github.com/lucylin/neural-benchmarks}}

In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}. The built-in cross-entropy loss functions in each framework were slightly different, which could potentially affect the number of iterations required for convergence.

Table~\ref{tab:hyperparams} summarizes the hyperparameters chosen for the experiments; there was no cross-validation done since the evaluative performance of the model trained is irrelevant.
\begin{table}\begin{center}
\begin{tabular}{cc}
\textbf{hyperparam} & \textbf{value} \\\hline
RNN type & \texttt{LSTM} \\
hidden size & $256$ \\
optimizer & \texttt{AdamOptimizer} \\
learning rate & $0.003$ \\
batch size & 25 \\
\end{tabular}
\caption{\label{tab:hyperparams}Hyperparameters used in the experiments}
\end{center}\end{table}

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. To determine if we should terminate, each full epoch of training (where all the training data is used) was followed by a predictive pass through the development data.

\section{Benchmarks}
\label{sec:benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns\footnote{We were not able to complete a convergence run for the Theano implementation in time for this paper. Based on the epoch-specific measurements and comparison of losses, Theano's performance on this task is likely to be comparable to TensorFlow's performance; we are confident that it would not surpass DyNet on this task.}. Overall, DyNet seems to converge the fastest of the three; its epoch time is dramatically shorter, but it also took more epochs to converge.

\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h21m 		& 35h25m 			& * \\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}

\subsection{Kernel function usage}
\label{subsec:kfunc}
Using NVidia's built-in GPU profiler, \verb!nvprof!, we first looked into which kernel function calls were most heavily utilized by profiling training on a single batch. As can be seen in Table~\ref{tab:pcttime}, a large portion of the runtime is spent on matrix multiplication (specifically, a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) for all frameworks. This is expected given that neural network operations are heavily dependent on matrix multiplies.

Unlike the other two frameworks, DyNet also spends a large amount of time in the \texttt{CUDA memset} function. This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.

\begin{table}
\centering
\begin{tabular}{c|rrr}
\textbf{Function} & DyNet &  TensorFlow & Theano \\  
\hline
matrix multiply & 29.64\%  &  68.87\% & 49.90\% \\
\hline
CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
\end{tabular}

\caption{\label{tab:pcttime}Percentage of runtime spent on the most common matrix multiply call used (a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) and on CUDA memset operations. These numbers are from training on a single batch.}
\end{table}

\begin{table*}
\centering
\begin{tabular}{c|rrr}
\textbf{Metric} 										& \textbf{DyNet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
\texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
\texttt{\detokenize{sm_efficiency}}					&		16.64\%	&		22.58\%	&		16.04\%		\\
\texttt{\detokenize{warp_efficiency}}					&		100.0\%	&		100.0\%	&		99.99\%		\\
\texttt{\detokenize{warp_nonpred_efficiency}}	&		100.0\%	&		99.95\%	&		99.91\%		\\
\texttt{\detokenize{global_hit_rate}}					&		2.88\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{local_hit_rate}}					&		0.00\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{dram_read_throughput}} (GB/s)		&		10.07	&		11.07	 	&		9.97 	\\
\texttt{\detokenize{dram_write_throughput}} (MB/s)	&		357.63&		411.99 &		503.54	\\
\end{tabular}

\caption{\label{tab:metrics} Various CUDA profiler metrics for the task of training a single batch. Descriptions of these metrics are in section \ref{subsec:cudaprof}.}
\end{table*}

\subsection{CUDA profiler metrics}
\label{subsec:cudaprof}

Given the high usage of the matrix multiply operation \verb!magma_lds128_sgemm_kernel()!, we
then looked into a variety of CUDA profiler metrics for this operation.

We collected data for the following metrics:
\begin{itemize}
\item \verb!achieved_occupancy! -- number of GPU warps used vs.\ total number of warps available.
\item \verb!sm_efficiency! -- how much of the runtime is spent performing actual computation.
\item \verb!warp_execution_efficiency! -- how efficient execution is within a warp; note that the number of registers per thread is likely a limiting factor for all implementations.
\item \verb!warp_nonpred_execution_efficiency! -- how efficient execution is within a warp on non-predicated (i.e. non-branching) instructions.
\item \verb!dram_read_throughput! -- read throughput from RAM to GPU.
\item \verb!dram_write_throughput! -- write throughput from GPU to RAM.
\end{itemize}

Table \ref{tab:metrics} summarizes the results on a single batch. We found that all of the measurements were within similar orders of magnitude; nothing at this level directly points to why DyNet performance was considerably faster than the other frameworks.

\section{Discussion}

Selecting a neural network framework involves several considerations. In this section, we describe how performance and various aspects of the programming experience may impact such a selection.

\subsection{Performance}

As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task.
DyNet seems to make certain optimizations that are advantageous in this setting, perhaps in the memory management choices it makes (as discussed in \ref{subsec:kfunc}). We also suspect (but could not empirically confirm based on metrics) that because DyNet dynamically creates computation graphs, it might be better able to adapt to different sentence lengths in a way that TensorFlow and Theano cannot.

Despite our performance findings, we note that the comparative performance between frameworks is likely dependent on the task, data set, choice of optimizer, and other parameters. Therefore, these results may not extrapolate to other settings, and if tuning training or prediction performance is a serious consideration, we recommend benchmarking on the specific task at hand.

\subsection{Programmer Experience}

While implementing language models in these frameworks, we found that it is also important to consider the development experience. No one framework wins or loses out; instead, there are several tradeoffs to consider. We describe some of our experiences below.

\subsubsection{Installation/dependencies}

Unlike a standard Python module which just requires a simple \texttt{pip install}, neural network frameworks are typically dependent on external fast low-level linear algebra libraries to perform efficiently. Therefore there is often linking or configuration required to wire things together.

\paragraph{DyNet} We encountered some difficulties in installing DyNet. The library failed to build with the recommended development version of Eigen, so we had to revert to an earlier commit. We also had to use Python 2 for the Python interface, though the documentation says it supports Python 3.

\paragraph{TensorFlow}
TensorFlow provides very extensive installation guidelines, and pip install takes care of installing it and its dependencies. However, one inconvenience is that it requires separate installations for CPU and GPU usage, which can cause issues if not using a virtual environment of some sort.

\paragraph{Theano} Our experience with Theano installation was relatively straightforward: pip install of relevant modules and specification of the BLAS install location.

\subsubsection{Ease of use}

\paragraph{DyNet}
DyNet includes built-in constructors for various NLP-relevant models, including several varieties of RNNs (e.g., tree LSTMs, encoder-decoders). This makes it easy for the programmer to build many common kinds of models. DyNet also allows for dynamic computation graphs, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths. While our neural language model used a static computation graph, DyNet made graph construction easy.

DyNet's primary drawback is the relative lack of documentation and support. The development team and userbase are both fairly small, so there is less information available to help users debug. There is some documentation available online, but it is incomplete.

\paragraph{TensorFlow} TensorFlow is relatively easy to use, as it comes with a myriad of tutorials and examples. TensorFlow's large userbase, polyvalence, and large amount of support from its creator (Google) have made it a very solid option for developers.

There is plenty of support for RNNs and many other standard models. However, writing the computation graph requires some amount of know-how, since there is a lot of TensorFlow-specific syntax, and debugging the symbolic graph can be difficult since there are no values. TensorFlow also sometimes requires a decent amount of boilerplate code; for example, the computation graph being statically constructed requires that sequences or mini-batches be padded, which can make RNN computation somewhat cumbersome.

\paragraph{Theano}
In contrast to DyNet and TensorFlow, Theano is ``lower-level'' in that it requires that the programmer define the shared variables propagation steps, and so on manually. Theano therefore offers a great deal of control and flexibility in model implementation, which is a benefit if one is implementing neural net layers not supported by other frameworks.

However, this also presents a steeper learning curve (despite the extensive documentation) and a much higher prototyping/development cost. We originally attempted to implement the Theano language model using just Theano and found the learning/debugging costs to be high. Because our language model consists of very standard neural net components, we instead switched to using Keras, which supplies an API with built-in RNN components implemented using Theano.

\section{Conclusion}
We implemented a simple language modeling task in three deep learning frameworks and measured their performance when run on a GPU. We found that DyNet converged faster than TensorFlow or Theano, but performance analysis on a fine-grained level showed that all frameworks used the underlying GPU resources with comparable efficiency. Through our experience implementing language models in these frameworks, we also found that efficiency is only one of several factors researchers should consider in choosing a deep learning framework. Users should also consider ease of installation and use, level of support available, and the suitability of the framework for their particular task.

\section*{Acknowledgments}
We want to thank Emily Furst for her help with GPU profiling (specifically, understanding the CUDA Profiler documentation).
\newpage
\bibliography{references}
\bibliographystyle{emnlp16}
\end{document}